A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-08-24 text/survival.Rmd
change some text to better describe the bayesian model, and change colors to avoid colorblindness problems in the kmcurves

**Figure 7: Comparison of different cancers, simulated and actual results
results.** A) To understand the risk model, we plot survival of four simulated
cohorts. Compare survival of a group consisting of randomly selected 30 year old
women (light blue) to an equal sized group of 30 year old women with cancer, k =
2 (light orange). To understand how the mix of ages alters the survival curve,
we compare a randomly selected group of people with ages uniformly distributed
from 30 to 70 (dark blue) and a randomly selected group with uniformly
distributed ages and cancer k = 2 (dark orange). B,C,D,E) We compare a randomly
selected group with correct age distribution for the given cancer (blue) to the
model predicted survival for the given average k (red) and the actual survival
for the patients in our dataset (CHOOSE COLOR HERE, NOT GREEN). As can be seen,
Stomach Cancer (STAD), Glioblastoma (GBM), and Pancreatic Cancer (PAAD) are all
very deadly. The modeled results fit the general trend for the actual patients.
Thyroid Cancer (THCA) on the other hand has very little if any effect on life
expectancy, in fact there may be a slight benefit which may be associated with a
bias in diagnosis towards patients who are more health conscious, or higher
socioeconomic status, with more access to care compared to random population
members.

@@

We found that cancer types, as expected, have a range of prognoses relative to
the general population. For example in **Figure 7B-E** we see three particularly
deadly cancers (Stomach:STAD, Glioblastoma:GBM, and Pancreatic:PAAD), and one
cancer where diagnosis apparently decreases risk relative to background
(Thyroid: THCA). Cancers with posterior probability for relative risk of less
than 1 such as THCA may be those where diagnosis is associated with patients who
have relatively healthy behaviors, are seen regularly by doctors, and comply
with health recommendations.

>>>>

**Figure 7: Comparison of different cancers, simulated and actual results
results.** A) To understand the risk model, we plot survival of four simulated
cohorts. Compare survival of a group consisting of randomly selected 30 year old
women (light blue) to an equal sized group of 30 year old women with cancer, k =
2 (light orange). To understand how the mix of ages alters the survival curve,
we compare a randomly selected group of people with ages uniformly distributed
from 30 to 70 (dark blue) and a randomly selected group with uniformly
distributed ages and cancer k = 2 (dark orange). B,C,D,E) We compare a randomly
selected group with correct age distribution for the given cancer (turqoise) to
the model predicted survival for the given average k (blue) and the actual
survival for the patients in our dataset (orange). As can
be seen, Stomach Cancer (STAD), Glioblastoma (GBM), and Pancreatic Cancer (PAAD)
are all very deadly. The modeled results fit the general trend for the actual
patients.  Thyroid Cancer (THCA) on the other hand has very little if any effect
on our estimates of life expectancy, in fact there may be a slight benefit which
may be associated with a bias in diagnosis towards patients who are more health
conscious, or higher socioeconomic status, with more access to care compared to
random population members.

@@

We found that cancer types, as expected, have a range of prognoses relative to
the general population. For example in **Figure 7B-E** we see three particularly
deadly cancers (Stomach:STAD, Glioblastoma:GBM, and Pancreatic:PAAD), and one
cancer where diagnosis apparently decreases risk relative to background
(Thyroid: THCA). Cancers with posterior probability for relative risk of less
than 1 such as THCA should be interpreted carefully. This Bayesian model is a
model for a state of information. The information that a person is diagnosed
with cancer may lead us to expect that they will live a shorter time than the
general population of matched age (k > 1), or a longer time than the general
population of matched age (k < 1). One mechanism for a shorter result is that
the cancer is aggressive and we can expect it to rapidly injure the body,
causing death. One explanation for longer results may be that the cancer is
relatively mild, and therefore diagnosis is potentially an informational signal
that the patient is health conscious, with the comparison group having more
people whose cancers go undiagnosed. It's important to note that the diagnosis
can increase our expectation of life relative to the comparison group, even if
it decreases the expectation of life of the individual relative to the
counterfactual where they did not have cancer.
